---
title: "BellaBeat"
author: "George.B"
date: "2022-11-20"
output:
  html_document: default
  pdf_document: default
---

# summary
## 1.The Business Project
* The task is to look out for trend on how people use the company smart devices and these insights can be     used by BellaBeat.

### Working with Rstudio.
Settings  the Rstudio and loading all necessary package to work with and then importing the BellaBeat dataset into my Rstudio.

```{r}
 library(tidyverse)
 library(dplyr)
 library(tidyr)
 library(here)
 library(ggplot2)
 library(colorspace)
 library(readr)
```

### importing the dataset.
```{r}

library(readr)
dailyActivity_merged <- read_csv("C:/Users/pc/Desktop/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
View(dailyActivity_merged)
sleepDay_merged <- read_csv("C:/Users/pc/Desktop/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
View(sleepDay_merged)
weightLogInfo_merged <- read_csv("C:/Users/pc/Desktop/Fitabase Data 4.12.16-5.12.16/weightLogInfo_merged.csv")
View(weightLogInfo_merged)
```

### Let do some quick summary check of the data columns.
```{r}
str(dailyActivity_merged)
colnames(dailyActivity_merged)
str(sleepDay_merged)
colnames(sleepDay_merged)
str(weightLogInfo_merged)
colnames(weightLogInfo_merged)
```

### Descriptive Analyses will now beging.
selecting and cleaning the data to wrangle.
In these case i will be selecting some vital columns from thr dailyActivity_merge dataset to work with.
```{r}
dailyActivity_merged2 <- dailyActivity_merged %>% 
  select(TotalDistance, VeryActiveDistance, ModeratelyActiveDistance, LightActiveDistance, Calories) %>% 
  drop_na() %>% 
  summarise(TotalDistance, VeryActiveDistance, ModeratelyActiveDistance, LightActiveDistance, Calories)
```

### visual representation
```{r}
dailyActive_merge_pattern <- dailyActivity_merged2 %>% 
  group_by(TotalDistance, Calories) %>%
  drop_na() %>% 
  summarise(TotalDistance, Calories) %>% 
  arrange(TotalDistance, Calories)
ggplot(data=dailyActive_merge_pattern)+
  geom_point(mapping=aes(x=TotalDistance, y=Calories, fill=Calories))+
  geom_smooth(mapping=aes(x=TotalDistance, y=Calories, fill=Calories))
```
The graph about shows that the more distance covered by a user the more calories they burn out of their body, and the lesser distance covered the smaller the calories the user burn out of they body.

### active records
here i am analysing the very active distance covered by each user against the time they use in covering that distance.
```{r}
dailyActivity_merged_veryactive <- dailyActivity_merged %>% 
  group_by(VeryActiveDistance, VeryActiveMinutes) %>% 
  drop_na() %>% 
  summarise(VeryActiveDistance, VeryActiveMinutes) %>% 
  arrange(VeryActiveDistance, VeryActiveMinutes)
ggplot(data=dailyActivity_merged_veryactive)+
  geom_point(mapping=aes(x=VeryActiveDistance, y=VeryActiveMinutes))
View(dailyActivity_merged_veryactive)
View(dailyActivity_merged_veryactive)
```

### Sleepday analyses
time to wrangle the second dataset on my workspace.
firstly i want to select, sort and summarise the particular columns to analyse.
```{r}
sleepday_01 <- sleepDay_merged %>% 
  separate(SleepDay, into = c('Date', 'Time'), sep= ' ')
print(sleepday_01)
```


```{r}
total_s_t <- sleepday_01%>% 
  group_by(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>% 
  drop_na() %>% 
  summarise(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed)
print(total_s_t)
```

```{r}
ggplot(data=sleepday_01)+
  geom_smooth(mapping=aes(x=TotalTimeInBed, y=TotalSleepRecords, group=1, color='TotalMinutesAsleep'))+
  geom_smooth(mapping=aes(x=TotalMinutesAsleep, y=TotalSleepRecords, group=2, color='TotalTimeInBed'))+
  labs(title = "BellaBeat Device: SleepingTime Vs. SleepRecords", subtitle = "Sleeping Pattern")
```
From the above presenation i analyse and visualised the relationship totalminutesasleep and totaltimeinbed by plottin the totalsleeprecords. accord to the graphical represantation, user spend more time sleeping in bed.

### weight analyses
here i will be analysing the weight of the users against their body mass index(BMI)

```{r}
weightactivity <- weightLogInfo_merged %>%
  select(WeightPounds, BMI, IsManualReport) %>% 
  group_by(WeightPounds, BMI, IsManualReport) %>% 
  drop_na() %>% 
  summarise(WeightPounds, BMI, IsManualReport) %>% 
  arrange(WeightPounds, BMI, IsManualReport)
print(weightactivity)
```

```{r}
ggplot(data=weightactivity)+
  geom_smooth(mapping = aes(x=WeightPounds, y=BMI, linetype=IsManualReport))+
  labs(title = "Bella Device: WeightPOunds Vs. BMI", subtitle = "Manual Report on Users")
```
this plot showcase a correlation between users weight and their body mass, The TRUE report conforms to the users weighting below 170pounds and having a body mass of less 30, While the FALSE report shows that user who weigh higher most surely have a higher body mass.

